Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue

### 2021-03-11 reports/rmarkdown/02_individual_reports.Rmd
Changes to run on Vildes computer

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  dataset: 1
---

@@

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
library(niceRplots)

@@

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data

>>>>

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  #pdf_document:
    #toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  dataset: 1
---

@@

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
library(scales)
remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
library(niceRplots)

@@

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
PATH <- "../.." #Path to the data

### 2021-03-11 reports/rmarkdown/03_integrated_datasets.Rmd
Changes to run on Vildes computer

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

>>>>

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatefeb22.csv"))
rownames(metadata) <- as.character(metadata$ID)

### 2021-03-11 reports/rmarkdown/manuscript/manuscript_template.Rmd
Changes to run on Vildes computer

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
library(igraph)
library(dplyr)
library(stringr)
library(rafalib)
library(scales)
library(dendextend)
library(edgeR)
remotes::install_github("czarnewski/niceRplots",force=T)
library(niceRplots)

@@

picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
```

```{r, echo = FALSE, eval = TRUE}
#Assign names to patient groups
big_groups <- c(
  L.inners=    "5",
  L.inners2=   "4",
  L.crispatus= "2",
  Gard=        "3",
  Gard.Prev=   "6",
  Gard.Prev2=  "8",
  Gard.Lacto=  "7",
  Patho=       "1"
)

@@

sizes <- as.matrix(clean_Ftest[o,o])
sizes[sizes > 7] <- 7
sizes <- sizes / max(sizes)
image( t(sizes [nrow(sizes):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(sizes)) ,
       labels = rownames(sizes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(sizes)) , par("usr")[3]-.02 ,
      labels = rownames(sizes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

@@

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

@@

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Monbilicus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardanerella",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,1,12))

shann <- vegan::diversity(t(x),index = c("shannon") )
simp <- vegan::diversity(t(x),index = c("simpson") )
invsimp <- vegan::diversity(t(x),index = c("invsimpson") )
divers <- as.matrix(rbind(shann,simp,invsimp) )
divers[is.infinite(divers)] <- 0
rownames(divers) <- c("Shannon","simpson","Inv Simpson")
barlist( data = divers,
         main = "",
         genes = c("Shannon","simpson","Inv Simpson"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

@@

taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
legend(x = par("usr")[c(2)],
       y = par("usr")[c(4)],title.adj = 0,
       legend = rownames(temp),xjust = 0,yjust = 1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
       pt.bg = taxa_pal)

@@

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16)

@@

@@

@@

layout(matrix(c(1,7,
                2,7,
                3,7,
                4,8,
                5,8,
                6,8) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))

@@

  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
  # legend(x = par("usr")[c(2)],
  #        y = par("usr")[c(4)],title.adj = 0,
  #        legend = rownames(temp),xjust = 0,yjust = 1,
  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
  #        pt.bg = taxa_pal)
}

@@

continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
continuous_metadata <- metadata[,continuous_parameters]
# continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )

@@

sizes <- as.matrix(clean_Ftest[o,o])
# sizes[sizes < 2] <- 0
sizes[sizes > 7] <- 7
sizes <- sizes / max(sizes)
sizes <- sizes * lower.tri(sizes)

mypar(1,1,mar=c(10,10,2,6))
plot(0,0, las=1,xlim=c(.5,ncol(sizes)+.5),ylim=c(.5,nrow(sizes)+.5),
       axes=F,ylab="",xlab="",main="",xaxs="i",yaxs="i", frame=T)
for(i in 1:ncol(sizes)){
  lines(c(i,i),c(nrow(sizes)+.5,0),col="grey95",lwd=.5,xpd=F)
}
for(i in 1:nrow(sizes) ){
  lines(c(0,ncol(sizes)+0.5),c(i,i),col="grey95",lwd=.5,xpd=F)
}
points(sort(rep(1:ncol(sizes),nrow(sizes))),
      rep(nrow(sizes):1,ncol(sizes)),
        col=colorRampPalette(c("grey95","orange3","firebrick","black"))(91)[cut( c(sizes) , breaks = seq(0,1,length.out = 92) )],pch=16 ,
      cex= sizes*1 + 0.2)
srt=90
text(1:ncol(sizes), par("usr")[3] - (par("usr")[4])/200, labels = colnames(sizes), srt = srt,
     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
     xpd = TRUE, cex=1)
text(par("usr")[1] , nrow(sizes):1 , labels = rownames(sizes), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=1)

@@

>>>>

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(igraph)
library(phyloseq)
library(vegan)
library(ggpubr)
library(rafalib)
library(scales)
library(dendextend)
library(edgeR)
remotes::install_github("czarnewski/niceRplots",force=T)
library(niceRplots)

@@

picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
picrust_KO_seq <- read.table("../../../data/KO_predicted.tsv",header = T,row.names = 1)
picrust_taxa <- read.table("../../../data/SeqID_to_taxa.csv",header = T,row.names = 1,sep = ",")
```

```{r, echo = FALSE, eval = TRUE}
#Assign names to patient groups
big_groups <- c(
  g2_L.iners=    "5",
  g3_L.iners2=   "4",
  g1_L.crispatus= "2",
  g5_Gard=        "3",
  g6_Gard.Prev=   "6",
  g7_Gard.Prev2=  "8",
  g4_Gard.Lacto=  "7",
  g8_Patho=       "1"
)

@@

sizes <- as.matrix(clean_Ftest[o,o])
sizes[sizes > 15] <- 15
sizes <- sizes / max(sizes)
image( t(sizes [nrow(sizes):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(sizes)) ,
       labels = rownames(sizes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(sizes)) , par("usr")[3]-.02 ,
      labels = rownames(sizes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(3,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

@@

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

@@

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Monbilicus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardanerella",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(2,5,1,12))

shann <- vegan::diversity(t(x),index = c("shannon") )
simp <- vegan::diversity(t(x),index = c("simpson") )
invsimp <- vegan::diversity(t(x),index = c("invsimpson") )
divers <- as.matrix(rbind(shann,simp,invsimp) )
divers[is.infinite(divers)] <- 0
rownames(divers) <- c("Shannon","simpson","Inv Simpson")
# barlist( data = divers,
#          main = "",
#          genes = c("Shannon","simpson","Inv Simpson"),
#          clustering = group_annotation,
#          draw_mean_lines=F,col = pal,srt=30)
  barlist( data = divers[,order(group_annotation)],
         main = "",
         genes = c("Shannon","simpson","Inv Simpson"),
         # clustering = group_annotation,
         draw_mean_lines=F,col = pal[factor(group_annotation[order(group_annotation)])])

@@

taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
legend(x = par("usr")[c(2)],
       y = par("usr")[c(4)],title.adj = 0,
       legend = rownames(temp),xjust = 0,yjust = 1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
       pt.bg = taxa_pal)
# 
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-1,
#         col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][sample_o],
#         pch=16,cex=.5,xpd=T)
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
#         col= c("grey","red")[as.numeric(metadata$HIVstatus)][sample_o],
#         pch=16,cex=.5,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
        pch=15,cex=.8,xpd=T)
  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-6,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.8,xpd=T)
  text(par("usr")[1],par("usr")[3]-6,labels = "HIVstatus",pos=2,cex=.8,xpd=T)

@@

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.005,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
        pch=16,xpd=T)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.01,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
        pch=16,xpd=T)

@@

```{r, echo = FALSE, eval = TRUE}
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
dim(picrust_taxa)
dim(picrust_KO_seq)

#Convert picrust to the same bacterial names in the manuscript
temp <- as.matrix(picrust_taxa)
temp[is.na(temp)] <- "NA"
sps <- apply( temp[,8:9] ,1,function(x) paste(x,collapse ="_"))
not_complete <- grepl("_NA",sps)
taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
taxa[!not_complete] <- sps[!not_complete]
taxa <- sub("_"," ",taxa)
length(taxa)
picrust_taxa$final_name <- taxa

dim(picrust_data)
dim(picrust_KO_seq)

common <- rownames(picrust_data)[rownames(picrust_data) %in% colnames(picrust_KO_seq)]
res <- t( t(as.matrix(picrust_data[common,])) %*% t(as.matrix(picrust_KO_seq[,common])) )
picrust_KO_seq <- picrust_KO_seq[,common]

picrust_KO_seq[1:100,1:100]

picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , as.character(picrust_taxa$Seq_ID)  ),"final_name"])
picrust_KO_seq <- as.matrix(picrust_KO_seq)

picrust_KO_seq[,1]

sub(" [(].*","",names(pvalues))[1]

ress <- sapply( sub(" [(].*","",names(pvalues)) , function(x){
  inn <- colnames(picrust_KO_seq) [ colnames(picrust_KO_seq) %in% names(gmt_list[[x]] ) ]
  temp <- picrust_KO_seq[,inn]
  temp <- temp[ order(rowMeans(temp),decreasing = T) ,  ]
  common_bacs <- rownames(SNN_bacteria)[rownames(SNN_bacteria) %in% rownames(temp)]
  temp <- rowsum(temp[common_bacs , ] , group = bac_communities [ rownames(SNN_bacteria) %in% common_bacs ] )
  return( temp )
} )

ress

mypar(1,2,mar=c(4,10,2,4))
x <- sapply(ress,function(x){
  rowMeans(x)
})
x [ is.nan(x) ] <- 0
x <- t(x)

sizes <- sapply(ress,function(x){
  return( rowSums( x > 0) / ncol(x) )
})
sizes [ is.nan(sizes) ] <- 0
sizes <- t(sizes)

plot( sort(rep(1:ncol(x),nrow(x))),
      rep(nrow(x):1,ncol(x)),
        col=colorRampPalette(c("grey95","firebrick"))(91)[cut( c(x) , breaks = seq(0,15,length.out = 92) )],pch=16 ,
      cex= sizes*2 + 0.2, ylim=c(0.5,nrow(x)+.5),yaxs="i",
      axes=F,xlab="",ylab="", frame=T)
srt=30
text(1:ncol(x), par("usr")[3] - (par("usr")[4])/200, labels = colnames(x), srt = srt,
     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
     xpd = TRUE, cex=.6)
text(par("usr")[1] - (par("usr")[2])/200, nrow(x):1 , labels = rownames(x), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=.6)

add_scale_legend(pal = colorRampPalette(c("grey95","firebrick"))(91),labels = c(min(x),round(max(x)) ))
add_size_legend(labels = c("0%","50%","100%"),max.cex = 2.2, min.cex = .2)

@@

image( t(top_TRX_counts[h$order,order(group_annotation[sample_use])][nrow(top_TRX_counts):1,]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .01,length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16,xpd=T)

points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
        pch=16,xpd=T)
points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.01,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
        pch=16,xpd=T)

@@

layout(matrix(c(1,7,
                2,7,
                3,8,
                4,8,
                5,8,
                6,8) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))

@@

  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
  # legend(x = par("usr")[c(2)],
  #        y = par("usr")[c(4)],title.adj = 0,
  #        legend = rownames(temp),xjust = 0,yjust = 1,
  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
  #        pt.bg = taxa_pal)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
        pch=15,cex=1,xpd=T)
  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-6,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=1,xpd=T)
  text(par("usr")[1],par("usr")[3]-6,labels = "HIVstatus",pos=2,cex=.8,xpd=T)

}

@@

continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_ng_mL_v3","Plasma_S_Estradiol_pg_mL_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
continuous_metadata <- metadata[,continuous_parameters]
# continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )

@@

sizes <- as.matrix(clean_Ftest[o,o])
# sizes[sizes < 2] <- 0
sizes[sizes > 10] <- 10
sizes <- sizes / max(sizes)
sizes <- sizes * lower.tri(sizes)

mypar(1,1,mar=c(10,10,2,6))
plot(0,0, las=1,xlim=c(.5,ncol(sizes)+.5),ylim=c(.5,nrow(sizes)+.5),
       axes=F,ylab="",xlab="",main="",xaxs="i",yaxs="i", frame=T)
for(i in 1:ncol(sizes)){
  lines(c(i,i),c(nrow(sizes)+.5,0),col="grey95",lwd=.5,xpd=F)
}
for(i in 1:nrow(sizes) ){
  lines(c(0,ncol(sizes)+0.5),c(i,i),col="grey95",lwd=.5,xpd=F)
}
points(sort(rep(1:ncol(sizes),nrow(sizes))),
      rep(nrow(sizes):1,ncol(sizes)),
        col=colorRampPalette(c("grey95","orange3","firebrick","red"))(91)[cut( c(sizes) , breaks = seq(0,1,length.out = 92) )],pch=16 ,
      cex= sizes*1 + 0.2)
srt=90
text(1:ncol(sizes), par("usr")[3] - (par("usr")[4])/200, labels = colnames(sizes), srt = srt,
     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
     xpd = TRUE, cex=1)
text(par("usr")[1] , nrow(sizes):1 , labels = rownames(sizes), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=1)

@@

## Figure S10

```{r, echo = FALSE, eval = TRUE, fig.show='hide'}
paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3
             '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region

n <- c("boston_r1", "boston_r2", "Sthlm")
ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

# remove controls and duplicates
ps_filt <- map(ps_list, ~ .x %>%
  subset_samples(., Include_final == "yes"))

# separate tissue from CVLv2 samples in the second Boston run
CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v2") 
tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == "v3")

# merge the lists
ps_types <- c(ps_filt, "CVLv2"=CVLv2, "tissue"=tissue) 
# Get ASV tables
seqtabs <- map(ps_types, ~data.frame(.x@otu_table@.Data))

# Get rarefaction curves
rare_runs <- map(seqtabs, ~ .x %>%
             rarecurve(., step = 20, label = FALSE)
)

ggplot_df_fun <- function(rare_obj, count) {
  # Add sample names 
  names(rare_obj) <- c(paste0("sample",sprintf("%03.0f", 1:length(rare_obj))))

  # Coerce data into "long" form.
  rare_df <- imap_dfr(rare_obj, ~ .x %>%
              as_tibble(., rownames = "Sample size") %>%
              mutate(Sample = .y) %>%
              mutate(`Sample size` = as.numeric(gsub("N","",`Sample size`)))
  )
  # Get nr. of counts in each sample 
  c <- formatC(count, format="f", big.mark = " ", digits=0)
  rare_df_col <- rare_df %>%
    dplyr::rename(ASV = "value") %>%
    group_by(Sample) %>% 
    nest() %>%
    mutate(Counts = map(data, ~max(.$`Sample size`))) %>%
    unnest(c(data, Counts)) %>%
    mutate(Counts = ifelse(Counts > count, paste0(">", c), paste0("<", c)))

  return(rare_df_col)
}

# Define some plot variables
count <- c(33000, 10000, 33000, 20000, 2500)
limit <- c(65000, 65000, 65000, 65000, 10000)
title <- c("Run B1. CVLv3", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B2. CVLv2", "Run B2. TISSUEv3")

# Extract data frame from the rarecurve object
rare_df_col<- map2(rare_runs, count, ~ggplot_df_fun(.x, .y))
# Create Plot 
rare_plot <- map(rare_df_col, 
              ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
              theme_bw() +
              scale_color_manual(values=c('tomato', alpha('black', 0.1))) + 
              geom_line(aes(color = Counts, group = Sample)) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.margin=unit(c(0,0,0.5,0),"cm"))
            )
raremax <- map(seqtabs, ~min(rowSums(.x)))
r_plot <- pmap(list(rare_plot, raremax, limit, title),
               ~ ..1 + xlim(c(0,..3)) +
               labs(title=paste0("Sequencing ", ..4)) +
               geom_vline(xintercept = ..2, 
                          linetype='dashed', 
                          color='slategray' )
)

# Save png
#imap(r_plot, ~ggsave(filename = paste0('Rarefaction_plot_', .y, '.png'), plot = .x, 
#        path = paste0("../../../results/"), 
#        width = 10, height = 5,)
#)

```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),
          r_plot[[3]] + theme(axis.title.x=element_blank()),
          r_plot[[4]] + theme(axis.title.x=element_blank()),
          r_plot[[5]],
          ncol = 1, 
          labels = c("A","B","C","D")
          )
```

**Supp. Figure 10. Rarefraction curves**. **A** Boston 1 sequencing run, **B** Sthlm sequencing run and **C** and **D** shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
